---
title: "Penguins"
output: html_notebook
---
# Load & Inspect Data
```{r}
penguins <- read.csv("penguins.csv")
head(penguins)
```

```{r}
library("dplyr")
```


```{r}
glimpse(penguins)
```
```{r}
str(penguins)
```

```{r}
is.numeric(penguins$Culmen.Depth..mm.)
```
```{r}
mean(penguins$Culmen.Depth..mm., na.rm=TRUE)
```
Check for duplicated rows.

```{r}
sum(duplicated(penguins))
```

```{r}
library(ggplot2)
ggplot(penguins, aes(x=Sex)) + geom_bar()
```
#### Look at misssing values 
```{r}
sapply(penguins, function(x) sum(is.na(x)))
```
```{r}
summary(penguins)
```
```{r}
list_na <- colnames(penguins)[ apply(penguins, 2, anyNA) ]
list_na
```
```{r}
n_distinct(penguins$Stage, na.rm = FALSE)
```

#### Omit missing values

```{r}
# penguins <- penguins %>% na.omit()
# dim(penguins)
```
Causes too much data to be lost
```{r}
class(penguins[3,11])
```
Drop comments column

```{r}
drop <- c("Comments", "Stage", "Region")
penguins = penguins[,!(names(penguins) %in% drop)]
```
Omit na values

```{r}
penguins <- penguins %>% na.omit()
dim(penguins)
```
```{r}
sapply(penguins, function(x) sum(is.na(x)))
```
#### All null values now removed

```{r}
penguins %>% count(Species)
```
```{r}
cor(penguins$Flipper.Length..mm, penguins$Culmen.Length..mm.)
```

```{r}
glimpse(penguins)
```
## Create visualisations for further EDA
```{r}
ggplot(penguins, aes(x=Flipper.Length..mm., fill=Species)) + 
  geom_histogram(bins=25, position = "dodge")
```


```{r}
ggplot(penguins, aes(x=Body.Mass..g., fill=Species)) + 
  geom_histogram(bins=25, position = "dodge")
```

```{r}
ggplot(penguins, aes(x=Flipper.Length..mm., fill=Species)) + 
  geom_boxplot() +
  coord_flip()
```

```{r}
ggplot(penguins, aes(x=Body.Mass..g., y=Flipper.Length..mm., color=Species)) + 
  geom_point()
```
```{r}
ggplot(penguins, aes(x=Culmen.Length..mm., y=Flipper.Length..mm., size=Body.Mass..g., color=Species)) + 
  geom_point() + 
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Flipper Length by Culmen Length, coloured by Species and mass shown by point size", x="Culmen Length (mm)", y="Flipper Length (mm)")
```
```{r}
ggplot(penguins, aes(x=Culmen.Length..mm., y=Flipper.Length..mm.)) +
  geom_point() + 
  facet_wrap(~ Species)
```

## Model Building
#### Start with basic tree models
```{r}
#install.packages("tidymodels")
library(tidymodels)
```
#### Simple Linear Model
```{r}
names(penguins)
```

which column only has one factor?
```{r}
which(sapply(penguins, function(x) (is.character(x) | is.factor(x)) & length(unique(x))<2))
```
column incldued in drop list

```{r}
#install.packages("caret")
library(caret)
set.seed(3456)

```

Get train and test data

```{r}
dt = sort(sample(nrow(penguins), nrow(penguins)*.7))
train<-penguins[dt,]
test<-penguins[-dt,]
```

```{r}
lm_model <- lm(Culmen.Length..mm. ~ ., data = train)
```

```{r}
lm_preds <- predict(lm_model, test)
```

```{r}
 # Random seed for reproducibility
set.seed(100)
# Create 10 folds of penguins
penguin_folds <- vfold_cv(train, v = 10)
```


build initial model
```{r}
tree_model <- decision_tree(tree_depth = 4, cost_complexity = 0.05) %>%
    set_mode("regression")
```

```{r}
tree_model <- model %>% set_engine("rpart")
```

```{r}
fits_cv <- fit_resamples(tree_model,
                         Culmen.Length..mm. ~ .,
                         resamples = penguin_folds,
                         metrics = metric_set(mae, rmse))
```


